---
title: "Course Project1 Part2"
author: "doyougnu"
date: "July 25, 2015"
output: html_document
---

##Overview

This brief report will explore the ToothGrowth base dataset with R, see C. I.
Bliss (1952) The Statistics of Bioassay. Academic Press for more information. It
will investigate the central questions of what variable/factors are
statistically significant given the data through the use of two-sample student
t-testing. This report finds that dose and dose administration method are both 
statistically significant factors affects the length of guinea pig teeth.

```{r dataPeek, echo = TRUE}
#load data
library(datasets)
library(ggplot2)
data("ToothGrowth")
df <- ToothGrowth

#lets look at the data
str(df)
head(df)

#do some quick plots
qplot(data = df, x = len, facets = supp ~ dose, color = supp)
```

Interesting it looks like data under label "VC" is right shifted compared to 
data under "OJ" for all levels of dose.

Lets look at the confidence intervals for each the data grouped by the supp 
variable, and by the dose variable

```{r confInt, echo = TRUE}
#perform unpaired t.test in R
tResults <- t.test(len ~ supp, data = df, paired = FALSE)

```

The above code executes a two sample t test on the ToothGrowth dataset. I use
the model statement format because it requires less code and is generally more
readable in my opinion. Read it as, the variable len, by the variable supp, in
the dataset df, unpaired t-test. One would generally use a paired t-test if the
experimental samples are identical through the course of the treatment. Or to 
relate it to the ToothGrowth dataset: The experiment uses 10 guinea pigs, and
administers 3 different Vitamin C dose levels **but splits** the guinea pigs
into two groups of Vitamin C application treatmenst. That is one group is "VC"
and the other is "OJ", hence it is unpaired because **the same guinea pig is not
undergoing both treament OJ and VC**.

Looking at the results of the t.test the 95% confidence intervals are 
`r tResults$conf.int` and the p-value was `r tResults$p.value` which is greater
than 0.05, hence we reject the alternate hypothesis (that the means of the two
samples is different), and accept the Null hypothesis (that the means of the two
samples is the same). Note that the p-value is very close to the alpha value, 
hence I would expect this to be statistically significant if more samples were
collected (doing so in light of this data would be a violation of scientific
standards in a study though - it is a well known method of p-value hacking)

Not lets investigate the statistical difference between the dose groups. 
Typically one would peform an ANOVA test to determine if any of the 3 dose 
groups are significantly different, as we are limited to students t tests I will
perform a t.test for each dose group combination. 

```{r doseTests, echo = TRUE}
#generate child datasets for VC
dose.0.5_VC <- df[df$dose == 0.5 & df$supp == "VC", ]
dose.1.0_VC <- df[df$dose == 1 & df$supp == "VC", ]
dose.2.0_VC <- df[df$dose == 2 & df$supp == "VC", ]

#generate child datasets for OJ
dose.0.5_OJ <- df[df$dose == 0.5 & df$supp == "OJ", ]
dose.1.0_OJ <- df[df$dose == 1 & df$supp == "OJ", ]
dose.2.0_OJ <- df[df$dose == 2 & df$supp == "OJ", ]

#t.test between 0.5 and 1.0
ttest_helper <- function(df1, df2, compVar = "len", paired = FALSE) {
    a <- t.test(df1[, compVar], df2[, compVar], paired = paired)
    a
}

#perform within supp t.tests
ttest_0.5_1.0_VC <- ttest_helper(dose.0.5_VC, dose.1.0_VC)
ttest_1.0_2.0_VC <- ttest_helper(dose.1.0_VC, dose.2.0_VC)
ttest_2.0_0.5_VC <- ttest_helper(dose.0.5_VC, dose.2.0_VC)

ttest_0.5_1.0_OJ <- ttest_helper(dose.0.5_OJ, dose.1.0_OJ)
ttest_1.0_2.0_OJ <- ttest_helper(dose.1.0_OJ, dose.2.0_OJ)
ttest_2.0_0.5_OJ <- ttest_helper(dose.0.5_OJ, dose.2.0_OJ)

#perform between supp groups t.tests
ttest_0.5_VC_OJ <- ttest_helper(dose.0.5_VC, dose.0.5_OJ) 
ttest_1.0_VC_OJ <- ttest_helper(dose.1.0_VC, dose.1.0_OJ) 
ttest_2.0_VC_OJ <- ttest_helper(dose.2.0_VC, dose.2.0_OJ) 

ttest_0.5_1.0_VC_OJ <- ttest_helper(dose.0.5_VC, dose.1.0_OJ) 
ttest_1.0_2.0_VC_OJ <- ttest_helper(dose.1.0_VC, dose.2.0_OJ) 
ttest_2.0_0.5_VC_OJ <- ttest_helper(dose.2.0_VC, dose.0.5_OJ) 
```

Now lets checkout the results. Notice that these t_tests were "paired" as the

##Results
```{r ttestResults, echo = TRUE, warning = FALSE}
#pack results in a container
ttest_list <- list(ttest_0.5_1.0_VC, ttest_1.0_2.0_VC, ttest_2.0_0.5_VC,
                ttest_0.5_1.0_OJ, ttest_1.0_2.0_OJ, ttest_2.0_0.5_OJ,
                ttest_0.5_VC_OJ, ttest_1.0_VC_OJ, ttest_2.0_VC_OJ,
                ttest_0.5_1.0_VC_OJ, ttest_1.0_2.0_VC_OJ, ttest_2.0_0.5_VC_OJ)

#I couldn't figure out how to do this programmatically, seems to be hard to
# in R
ttest_list_names <-
    c("ttest_0.5_1.0_VC", "ttest_1.0_2.0_VC", "ttest_2.0_0.5_VC",
      "ttest_0.5_1.0_OJ", "ttest_1.0_2.0_OJ", "ttest_2.0_0.5_OJ",
      "ttest_0.5_VC_OJ", "ttest_1.0_VC_OJ", "ttest_2.0_VC_OJ",
      "ttest_0.5_1.0_VC_OJ", "ttest_1.0_2.0_VC_OJ", "ttest_2.0_0.5_VC_OJ")

#extract pvalues and confidence intervals
ttest_list_pvals <- sort(unlist(lapply(ttest_list, function(x) {x[3][1]})))
ttest_list_confs <- lapply(ttest_list, function(x) {x[4]})

#Make data table
library(data.table)
dt <- data.table(T_Test = 0, P_Value = rep(0, 12), Conf_Int1 = 0, Conf_Int2 = 0)
dt[,1] <- ttest_list_names
dt[,2] <- ttest_list_pvals
dt[,3] <- ttest_list_confs[[1]]
dt[,4] <- ttest_list_confs[[2]]

#convert to real table
library(knitr)
kable(dt)
```


##Conclusions
Starting from the most significant p-value, we can see that every level of dose,
when compared to each other level is statistically significant - that is to say
that we reject the Null Hypothesis (that the means of each group are equal) for
the Alternate Hypothesis (that the means of each group are not equal) for all
but the last three rows in the table. In sum, Dose is a statistically
significant variable in this experiment when convolved with the "supp" variable,
and when deconvloved with the "supp" variable. The only time dose is not
statistically significant is in the 1.0 dose VC, 2.0 dose OJ treatment condition
and 2.0_VC, 0.5 OJ treatment condtion. It is important to note that when OJ and
VC were compared globally they were not statistically different. When compared
within each dose treatment group **they are** statistically different.